Eigenvector centrality signifies antidepressants as a central hub of the co-medication network

### 2024-08-14 docs/abstract/_abstract-full.qmd
Incorporate revision suggested by Spyros

**Background:**
The implementation of network analysis in medicine and epidemiology started to gain popularity. A network-based approach captures the relationships of the building blocks of a complex system. Drug co-prescription is a complex issue presenting with one-to-many relationships that may change over time, especially also in mental disorders.

**Objectives:**  
Evaluate the temporal relationships of drug treatment for mood disorders with other drug therapies in a pharmacy dispensing database.

**Methods:**
Exploratory temporal network analysis was performed using the University of Groningen prescription database, IADB.nl. A closed cohort with at least one prescription of anxiolytics/antidepressants dispensed anywhere from 2018 to 2022 was extracted. The drug prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification, where psychopharmaca was evaluated at ATC level 4, drugs for the nervous system at ATC level 3, and other medications at ATC level 1, distinguished into a total of 24 therapeutic medication classes. A drug prescription network was constructed as an undirected graph to capture dose-adjusted co-prescription. Eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of prescriptions ($n_c$), number of patients, claim-to-patient ratio, and eigenvector centrality were extracted as weekly-aggregated time-series data. Singular spectrum analysis was performed to decompose the data into its constituting trend, harmonics, and noise components.

>>>>

**Background:**
Networks have been used in different scientific fields to model complex systems. In medicine and epidemiology, drug prescription networks (DPN) have been proposed as a modeling tool for studying the temporal dynamics of medication distribution and consumption within a population.

**Objectives:**  
Evaluate the structural properties of temporal DPNs that are constructed from the observed daily co-prescriptions between treatments for mental health problems and other drug therapies in a pharmacy dispensing database.

**Methods:**
For  exploratory time series analysis, and network construction, we utilized the IADB.nl database that collects prescriptions from 128 pharmacies in the Netherlands. A closed cohort with at least one prescription of anxiolytics/antidepressants dispensed anywhere from 2018 to 2022 was extracted. The drug prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification, where psychopharmaca was evaluated at ATC level 4, drugs for the nervous system at ATC level 3, and other medications at ATC level 1, distinguished into a total of 24 therapeutic medication classes. Symmetric daily dose-adjusted co-prescriptions were used to construct time-varying DPNs as undirected graphs. Subsequently, eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of prescriptions ($n_c$), number of patients, claim-to-patient ratio, and eigenvector centrality were extracted as weekly-aggregated time-series data. Singular spectrum analysis was performed to decompose the data into its constituting trend, harmonics, and noise components.
